---
title: "Data Cleaning Process - All"
author: "MSBA-SA Capstone"
date: "Spring 2024"
output: html_document
---

The file contains information relevant to the data cleaning throughout the data analysis process. Throughout the code, there are lines to uncomment, using `write.csv()` to save the data to the working directory. The team saved the cleaned data files and then loaded the files into the environment to save time throughout the process. 

## <span style= "color:#00843D" > DATA IMPORTATION & SET UP </span>

```{r packages, warnings=FALSE, error=FALSE, message=FALSE}
library(tidyverse)
library(DescTools)
library(janitor)
library(ggplot2)
library(jpeg)
library(chron)
library(lubridate)
library(anytime)
library(ggpubr)
library(devtools) 
#install_github("AppliedDataSciencePartners/xgboostExplainer")
library(xgboost) # Load XGBoost
library(caret) # Load Caret
library(xgboostExplainer) # Load XGboost Explainer
library(pROC) # Load proc
library(SHAPforxgboost) # Load shap for XGBoost
library(plyr)
library(Metrics)

# FOR WEB SCRAPING
library(httr) 
library(rvest) 
library(polite) 
library(jsonlite) 
```

```{r nd colors}
# setting ND colors 
nd_navy <- "#0C2340"
nd_green <- "#00843D"
nd_gold <- "#C99700"
```

```{r creating functions}
calc_num_zeros_by_col <- function(df_title) {
  cols_vec <- colnames(df_title)
  zeros_count_vec <- rep("", length=length(cols_vec))
  zeros_test_vec <- rep("", length=length(cols_vec))

  for (i in 1:length(cols_vec)){
    zeros_count_vec[i] <- sum(df_title[,i] == 0)
    zeros_test_vec[i] <- sum(df_title[,i] == 0 | df_title[,i] == "0")
  }

  zero_count_df <- cbind.data.frame(cols_vec, as.numeric(zeros_count_vec))
  names(zero_count_df) <- c("column_name", "number_zeros")

  # Append df_title to the dataframe name
  assign(paste0("zero_count_df_", deparse(substitute(df_title))), zero_count_df,
         envir = .GlobalEnv)
}
```

```{r}
lax0_data <- read.csv("LAXdata0.csv") #reading in data
lax1_data <- read.csv("LAXdata1.csv") #reading in data
soc0_data <- read.csv("SOCdata0.csv") #reading in data
soc1_data <- read.csv("SOCdata1.csv") #reading in data
```


## <span style= "color:#00843D" > DATA CLEANING - INITIAL </span>
```{r}
cutoff_val <- .95
####################### LAX 0 ####################### 
lax0_data <- clean_names(lax0_data) #cleaning names to snakecase
#dim(lax0_data) #check dimensions
#create a cutoff to use for dropping a column if more than that percent are zero
drop_var_cutoff_lax0 <- dim(lax0_data)[1]*cutoff_val

####################### LAX 1 #######################
lax1_data <- clean_names(lax1_data)  #cleaning names to snakecase
#dim(lax1_data) #check dimensions
drop_var_cutoff_lax1 <- dim(lax1_data)[1]*cutoff_val #lax 1 drop cutoff

####################### SOC 0 #######################
soc0_data <- clean_names(soc0_data) #cleaning names to snakecase
#dim(soc0_data) #check dimensions
drop_var_cutoff_soc0 <- dim(soc0_data)[1]*cutoff_val #soc 0 drop cutoff

####################### SOC 1 #######################
soc1_data <- clean_names(soc1_data) #cleaning names to snakecase
#dim(soc1_data) #check dimensions
drop_var_cutoff_soc1 <- dim(soc1_data)[1]*cutoff_val #soc 0 drop cutoff
```


```{r}
cols_to_remove_vec <- c("month_name", "day_name", "x",
                        "team_name", "total_activities", "athlete_weight", 
                        "player_name", "jersey")

####################### LAX 0 #######################
calc_num_zeros_by_col(lax0_data)
#zero_count_df_lax0_data
miss_over_cutoff_perc_df_lax0_data <- zero_count_df_lax0_data %>% 
  filter(number_zeros >= drop_var_cutoff_lax0)
lax0_data <- lax0_data %>%
  select(-one_of(miss_over_cutoff_perc_df_lax0_data$column_name)) %>%
  select(!all_of(cols_to_remove_vec)) %>%
  select(-contains("heart")) 

# changing position name to a numeric factor
lax0_data$position_name <- as.numeric(revalue(lax0_data$position_name, 
        c("Attack"=1, "D-Mid"=2, "Defense"=3, 
          "FOS"=4, "LSM"=5, "Midfield"=6)))

# adding a column for a "unique session" --> player, date, start time
lax0_data$unique_session <- paste(lax0_data$name_id, 
                                  lax0_data$date, 
                                  lax0_data$unix_start_time, 
                                  sep="-")
```

```{r}
####################### LAX 1 #######################
calc_num_zeros_by_col(lax1_data)
#zero_count_df_lax1_data
miss_over_cutoff_perc_df_lax1_data <- zero_count_df_lax1_data %>%
  filter(number_zeros >= drop_var_cutoff_lax1)
lax1_data <- lax1_data %>%
  select(-one_of(miss_over_cutoff_perc_df_lax1_data$column_name)) %>%
  select(-contains(cols_to_remove_vec))

# changing position name to a numeric factor
pos_options <- unique(lax1_data$position_name)
lax1_data$position_name <- as.numeric(revalue(lax1_data$position_name, 
        c("Midfielder"=1, "Midfielder - Attacking"=2, "Attacker"=3, 
          "Defender"=4, "Midfielder - Defensive"=5)))

# adding a column for a "unique session" --> player, date, start time
lax1_data$unique_session <- paste(lax1_data$name_id, 
                                  lax1_data$date, 
                                  lax1_data$start_time, 
                                  sep="-")
```

```{r}
####################### SOC 0 #######################
calc_num_zeros_by_col(soc0_data)
#zero_count_df_soc1_data
miss_over_cutoff_perc_df_soc0_data <- zero_count_df_soc0_data %>%
  filter(number_zeros >= drop_var_cutoff_soc0)
soc0_data <- soc0_data %>%
  select(-one_of(miss_over_cutoff_perc_df_soc0_data$column_name)) %>%
  select(-contains(cols_to_remove_vec))

# changing position name to a numeric factor
pos_options <- unique(soc0_data$position_name)
soc0_data$position_name <- as.numeric(revalue(soc0_data$position_name, 
        c("Center Back"=4,
          "Center Mid"=2,
          "Forward"=1,
          "Outside Mid"=6, 
          "Outside Back"=5,
          "GoalKeeper"=0, 
          "6/8"=3,#midfield
          "2/3"=5,#outside back
          "7/11"=1,  #outside forward
          "9"  =1, #striker
          "1" = 0, #GK 
          "4/5" =4,
          "Goalkeeper"=0, 
          "Defensive Line"=4, 
          "Defensive Back"   =4,
          "Central Midfielder"=3, 
          "Defender"=4,
          "Missing Position"=7, 
          "Fullback"=5, 
          "Central Defender"=4, 
          "Winger"=6)))

# adding a column for a "unique session" --> player, date, start time
soc0_data$unique_session <- paste(soc0_data$name_id, 
                                  soc0_data$date, 
                                  soc0_data$start_time, 
                                  sep="-")
```

```{r}
####################### SOC 1 #######################
calc_num_zeros_by_col(soc1_data)
#zero_count_df_soc1_data
miss_over_cutoff_perc_df_soc1_data <- zero_count_df_soc1_data %>%
  filter(number_zeros >= drop_var_cutoff_soc1)
soc1_data <- soc1_data %>%
  select(-one_of(miss_over_cutoff_perc_df_soc1_data$column_name)) %>%
  select(-contains(cols_to_remove_vec)) %>%
  select(-contains(c("average_high_speed_distance",
  "high_speed_running_distance")))


# FORWARD 1, MIDFIELD 2, BACK 3, Utility 4, WINGER 5, GK 0
soc1_data$position_name <- as.numeric(revalue(soc1_data$position_name, 
        c("Forward"=1, "Midfield"=2, "Center Back" = 3,
          "Striker"=1, "Keeper" = 0, "Outside Back" = 4,
          "Center Midfield"=2, "Defender"=3, "Midfielder" = 2, 
          "Goal Keeper"=0, "Defensive Midfielder"=2, "Central Midfielder"=2,
          "Attacking Midfielder"=2, "Centre Back"=3, "Utility"=4, 
          "Wing"=5, "Back"=3, "Wide"=5)))

# adding a column for a "unique session" --> player, date, start time
soc1_data$unique_session <- paste(soc1_data$name_id, 
                                  soc1_data$date, 
                                  soc1_data$start_time, 
                                  sep="-")
```


```{r}
lax0_data$location <- as.factor(ifelse(lax0_data$total_distance == 0,
                                       "indoor", "outdoor"))
lax1_data$location <- as.factor(ifelse(lax1_data$total_distance == 0,
                                       "indoor", "outdoor"))
soc0_data$location <- as.factor(ifelse(soc0_data$total_distance == 0,
                                       "indoor", "outdoor"))
soc1_data$location <- as.factor(ifelse(soc1_data$total_distance == 0,
                                       "indoor", "outdoor"))

not_include_indoor_vec <- c("duration", "distance", "meta", "total_effort", 
                            "heart_rate", "hr", "velocity", "acceleration",
                            "deceleration", "metre", "meterage",
                            "exertion_index","max_vel_max", 
                            "footstrikes", "running_series_count", 
                            "running_imbalance")

lax0_indoor_full_df <- lax0_data %>%
  filter(location=="indoor") %>%
  select(-contains(not_include_indoor_vec))

lax1_indoor_full_df <- lax1_data %>%
  filter(location=="indoor") %>%
  select(-contains(not_include_indoor_vec))

soc0_indoor_full_df <- soc0_data %>%
  filter(location=="indoor") %>%
  select(-contains(not_include_indoor_vec))

soc1_indoor_full_df <- soc1_data %>%
  filter(location=="indoor") %>%
  select(-contains(not_include_indoor_vec))
```



```{r}
# testing_names_vec <- lax0_data %>% select_if(is.character) %>%
#   select(contains(c("time", "duration"))) %>% names()
  
# for (i in seq_along(testing_names_vec)) {
#   col_name <- testing_names_vec[i]
#   
#   lax0_data[[col_name]] <- chron(times = lax0_data[[col_name]])
#   lax0_data[[col_name]] <- hms(lax0_data[[col_name]])
#   lax0_data[[col_name]] <- hour(lax0_data[[col_name]])*60 +
#     minute(lax0_data[[col_name]])
# }


#################################### LAx 0 #############################
dur_chr_df <- lax0_data %>% 
  select(contains(c("duration", "time")))%>% 
  select_if(is.character)

for (q in seq_along(names(dur_chr_df))){
  col_vec <- dur_chr_df[, q]
  col_name <- names(dur_chr_df[q])

  col_vec <- str_replace_all(col_vec, pattern="[[:punct:]]", replacement = "")
  
  for (r in 1:length(col_vec)){
   hrs_val <- as.numeric(substr(col_vec[r], 1, 2))
   min_val <- as.numeric(substr(col_vec[r], 3, 4))
   tot_min <- as.numeric((hrs_val*60) + min_val) 
   col_vec[r] <- tot_min
  }
  
  dur_chr_df[, q] <- as.numeric(col_vec)
  lax0_data[[col_name]] <- dur_chr_df[, q]
}




#################################### LAX 1 #############################
dur_chr_df <- lax1_data %>%
  select(contains(c("duration", "time")))%>%
  select_if(is.character)

for (q in seq_along(names(dur_chr_df))){
  col_vec <- dur_chr_df[, q]
  col_name <- names(dur_chr_df[q])

  col_vec <- str_replace_all(col_vec, pattern="[[:punct:]]", replacement = "")

  for (r in 1:length(col_vec)){
   hrs_val <- as.numeric(substr(col_vec[r], 1, 2))
   min_val <- as.numeric(substr(col_vec[r], 3, 4))
   tot_min <- as.numeric((hrs_val*60) + min_val)
   col_vec[r] <- tot_min
  }

  dur_chr_df[, q] <- as.numeric(col_vec)
  lax1_data[[col_name]] <- dur_chr_df[, q]
}






#################################### SOC 0 #############################
dur_chr_df <- soc0_data %>%
  select(contains(c("duration", "time"))) %>%
  select_if(is.character)

for (q in seq_along(names(dur_chr_df))){
  col_vec <- dur_chr_df[, q]
  col_name <- names(dur_chr_df[q])

  col_vec <- str_replace_all(col_vec, pattern="[[:punct:]]", replacement = "")

  for (r in 1:length(col_vec)){
   hrs_val <- as.numeric(substr(col_vec[r], 1, 2))
   min_val <- as.numeric(substr(col_vec[r], 3, 4))
   tot_min <- as.numeric((hrs_val*60) + min_val)
   col_vec[r] <- tot_min
  }

  dur_chr_df[, q] <- as.numeric(col_vec)
  soc0_data[[col_name]] <- dur_chr_df[, q]
}




#################################### SOC 1 #############################
dur_chr_df <- soc1_data %>%
  select(contains(c("duration", "time")))%>%
  select_if(is.character)

for (q in seq_along(names(dur_chr_df))){
  col_vec <- dur_chr_df[, q]
  col_name <- names(dur_chr_df[q])

  col_vec <- str_replace_all(col_vec, pattern="[[:punct:]]", replacement = "")

  for (r in 1:length(col_vec)){
   hrs_val <- as.numeric(substr(col_vec[r], 1, 2))
   min_val <- as.numeric(substr(col_vec[r], 3, 4))
   tot_min <- as.numeric((hrs_val*60) + min_val)
   col_vec[r] <- tot_min
  }

  dur_chr_df[, q] <- as.numeric(col_vec)
  soc1_data[[col_name]] <- dur_chr_df[, q]
}
```


### <span style= "color:#00843D" > ADDING GAME/TRAINING COLUMN </span>
retrieving information from scraped game data
```{r}
# need to first run the scraping sports schedule file to save the following data in the environment
#source("scraping_sport_schedules.r")
lax0_game_dates <- read.csv(".//Exported_CSVs//lax0_game_dates.csv")$x
lax1_game_dates <- read.csv(".//Exported_CSVs//lax1_game_dates.csv")$x
soc0_game_dates <- read.csv(".//Exported_CSVs//soc0_game_dates.csv")$x
soc1_game_dates <- read.csv(".//Exported_CSVs//soc1_game_dates.csv")$x
```


```{r}

####################### LAX 0 #######################
lax0_data$activity_type <- as.factor(ifelse(lax0_data$date %in% lax0_game_dates, 
                                  "game", "training"))
lax0_data$activity_type_binary <- as.numeric(ifelse(lax0_data$date %in% lax0_game_dates, 
                                  1, 0))

####################### LAX 1 #######################
lax1_data$activity_type <- as.factor(ifelse(lax1_data$date %in% lax1_game_dates, 
                                  "game", "training"))
lax1_data$activity_type_binary <- as.numeric(ifelse(lax1_data$date %in% lax1_game_dates, 
                                  1, 0))

####################### SOC 0 #######################
soc0_data$activity_type <- as.factor(ifelse(soc0_data$date %in% soc0_game_dates, 
                                  "game", "training"))
soc0_data$activity_type_binary <- as.numeric(ifelse(soc0_data$date %in% soc0_game_dates, 
                                  1, 0))

####################### SOC 1 #######################
soc1_data$activity_type <- as.factor(ifelse(soc1_data$date %in% soc1_game_dates, 
                                  "game", "training"))
soc1_data$activity_type_binary <- as.numeric(ifelse(soc1_data$date %in% soc1_game_dates, 
                                  1, 0))
```

Uncomment the following lines to save the cleaned data as a csv
```{r}
# write.csv(lax0_data,".\\Exported_CSVs\\lax0_data_cleaned.csv",row.names=FALSE)
# 
# write.csv(lax1_data,".\\Exported_CSVs\\lax1_data_cleaned.csv",row.names=FALSE)
# 
# write.csv(soc0_data,".\\Exported_CSVs\\soc0_data_cleaned.csv",row.names=FALSE)
# 
# write.csv(soc1_data,".\\Exported_CSVs\\soc1_data_cleaned.csv",row.names=FALSE)

#loading in data to confirm
# cleaned_data_lax0 <- read.csv(".\\Exported_CSVs\\lax0_data_cleaned.csv") 
```


### <span style= "color:#00843D" > COMBINING SOCCER DATA TOGETHER INTO ONE DATASET FOR MODELING </span>

##### LOADING CLEAN DATA 
```{r}
soc0_data <- read.csv(".\\Exported_CSVs\\soc0_data_cleaned_up_pos.csv") 
soc1_data <- read.csv(".\\Exported_CSVs\\soc1_data_cleaned_up_pos.csv") 
```

```{r}
not_include_indoor_vec <- c("meta", "total_effort", 
                            "heart_rate", "hr", "velocity", "acceleration",
                            "deceleration", "metre", "meterage",
                            "exertion_index","max_vel_max", 
                            "footstrikes", "running_series_count", 
                            "running_imbalance"#," distance", "duration"
                            )

length(names(soc0_data)) #826 columns
length(names(soc1_data)) #701 columns


# indoor_diff_01 <- setdiff(names(soc0_data), names(soc1_data))
# indoor_diff_10 <- setdiff(names(soc1_data), names(soc0_data))
# full_indoor_diff <- c(indoor_diff_01, indoor_diff_10)

# testing_vec <- c("velocity_band_6_distance_band_1_total_efforts", 
#                  "velocity_band_7_distance_band_1_total_efforts", 
#                  "velocity_band_7_recovery_band_2_total_efforts" , 
#                  "velocity_b2_average_effort_distance_gen_2_set_2",
#                  "velocity_b2_total_efforts_gen_2_set_2")

#names(soc1_data) <- str_replace_all(names(soc1_data), "_b([0-9])", "_band_\\1")

# diff01 <- setdiff(names(soc0_data), names(soc1_data))
# diff10 <- setdiff(names(soc1_data), names(soc0_data))
# 
# full_diff <- c(diff01, diff10)
# length(full_diff) #245 originally

cols_exclude <- c("mii", "metabolic", "heart_rate", "_set_2")
soc0_data <- soc0_data %>% 
  select(!contains(cols_exclude))

soc1_data <- soc1_data %>% 
  select(!contains(cols_exclude))

soc1_data <- soc1_data %>%
  dplyr::rename(
    velocity_band_1_total_distance = velocity_band_1_0_45mph_total_distance,
    velocity_band_2_total_distance = velocity_band_2_45_4_05mph_total_distance,
    velocity_band_3_total_distance = velocity_band_3_4_05_7_14mph_total_distance,
    velocity_band_4_total_distance = velocity_band_4_7_14_9_64mph_total_distance,
    velocity_band_5_total_distance = velocity_band_5_9_64_12_12mph_total_distance,
    velocity_band_6_total_distance = velocity_band_6_12_12_14_61mph_total_distance,
    velocity_band_7_total_distance = velocity_band_7_14_61_mph_total_distance, 
    high_speed_distance = total_high_speed_distance, 
    pl_hi=pl_hi_session
  )

soc0_data$med_and_hi_ima_r <- soc0_data$ima_co_d_right_medium_1_0+soc0_data$ima_co_d_right_high_1_0
soc0_data$med_and_hi_ima_l <- soc0_data$ima_co_d_left_medium_1_0+soc0_data$ima_co_d_left_high_1_0
#names(soc1_data)

fr0 <- soc0_data %>% select(contains("ima_free_running")) %>% names()
fr1 <- soc1_data %>% select(contains("ima_free_running")) %>% names()

free_run_exclude <- setdiff(fr0, fr1)

soc0_data <- soc0_data %>% select(-contains(c(free_run_exclude)))
```

```{r}
diff01 <- setdiff(names(soc0_data), names(soc1_data))
diff10 <- setdiff(names(soc1_data), names(soc0_data))


full_diff <- c(diff01, diff10)
length(full_diff) #245 originally

soc1_indoor_full_df <- soc1_data %>%
  select(-contains(not_include_indoor_vec))

soc0_indoor_full_df <- soc0_data %>%
  select(-contains(not_include_indoor_vec))


# should only look at differences for variables available indoors
indoor_diff_01 <- setdiff(names(soc0_indoor_full_df), names(soc1_indoor_full_df))
indoor_diff_10 <- setdiff(names(soc1_indoor_full_df), names(soc0_indoor_full_df))
full_indoor_diff <- c(indoor_diff_01, indoor_diff_10)
```


```{r}
exclude_to_combine <- full_indoor_diff
# soc0_indoor_full_df <- soc0_indoor_full_df %>% select(-contains(c(exclude_to_combine)))
# soc1_indoor_full_df <- soc1_indoor_full_df 

soc1_indoor_full_df <- soc1_indoor_full_df %>%
  #select(-contains(not_include_indoor_vec)) %>% 
  select(-contains(c(exclude_to_combine)))

soc0_indoor_full_df <- soc0_indoor_full_df %>%
  #select(-contains(not_include_indoor_vec)) %>% 
  select(-contains(c(exclude_to_combine)))

# should only look at differences for variables available indoors
indoor_diff_01 <- setdiff(names(soc0_indoor_full_df), names(soc1_indoor_full_df))
indoor_diff_10 <- setdiff(names(soc1_indoor_full_df), names(soc0_indoor_full_df))
full_indoor_diff <- c(indoor_diff_01, indoor_diff_10)


# confirm names are sorted alphabetically
# names(soc0_mod_sort)
# names(soc1_mod_sort)
dur_cols_remove0 <- soc0_indoor_full_df %>% 
  select(contains("duration")) %>% names()
dur_cols_remove1 <- soc1_indoor_full_df %>% 
  select(contains("duration")) %>% names()
dur_cols_remove <- c(dur_cols_remove0, dur_cols_remove1)
dur_cols_remove <- dur_cols_remove[dur_cols_remove != "total_duration"]  

dist_cols_remove0 <- soc0_indoor_full_df %>% 
  select(contains("distance")) %>% names()
dist_cols_remove1 <- soc1_indoor_full_df %>% 
  select(contains("distance")) %>% names()
dist_cols_remove <- c(dist_cols_remove0, dist_cols_remove1)
dist_cols_remove <- dist_cols_remove[dist_cols_remove != "high_speed_distance"]  

dist_dur_cols_remove <- c(dur_cols_remove, dist_cols_remove)

soc0_indoor_full_df <- soc0_indoor_full_df %>%
  select(-contains(dist_dur_cols_remove))

soc1_indoor_full_df <- soc1_indoor_full_df %>%
  select(-contains(dist_dur_cols_remove)) 
#soc_full_mod_df$dist_dur_cols_remove

soc0_indoor_full_df$team_gender <- 0
soc1_indoor_full_df$team_gender <- 1

soc0_mod_sort <- soc0_indoor_full_df %>% select(sort(names(.)))
soc1_mod_sort <- soc1_indoor_full_df %>% select(sort(names(.)))

dim(soc0_mod_sort)
dim(soc1_mod_sort) 



#binding dataframes
soc_full_mod_df <- rbind.data.frame(soc0_mod_sort, soc1_mod_sort)
```

```{r}
# write.csv(soc_full_mod_df,".\\Exported_CSVs\\comb_soc_data_cleaned.csv",row.names=FALSE)
```


```{r}
players <- unique(soc_full_mod_df$name_id)

z_prep_mod_df <- soc_full_mod_df

full_df <- as.data.frame(matrix(ncol=dim(z_prep_mod_df)[2]))
names(full_df) <- names(z_prep_mod_df)

for(i in 1:length(players)){
  player_df <- z_prep_mod_df[z_prep_mod_df$name_id == players[i],]
  
  df_non_num <- player_df %>% select_if(is.character)
  df_num <- player_df %>% select_if(is.numeric)
  
  df_num <- scale(df_num)
  combined <- cbind(df_non_num, df_num)
  full_df <- rbind(full_df, combined)
}

full_df <- full_df[-1,]
names(full_df) <- paste0(names(full_df), "_z") 

z_score_df <- full_df %>% select(-contains(c("name_id")))
```



```{r}
comb_z_df <- merge(soc_full_mod_df, z_score_df, 
                   by.x = "unique_session", by.y = "unique_session_z",
                   all.x=TRUE)
#head(comb_z_df)
```



```{r warning=FALSE}
num_col_names <- soc_full_mod_df %>% select_if(is.numeric) %>% names()
num_col_names_indoor <- num_col_names[num_col_names %in% names(soc_full_mod_df)]

#to_remove_vec <- c("unix", "position_name_z")

test_df <- soc_full_mod_df %>%
  select(all_of(c("unique_session","name_id", "high_speed_distance", num_col_names_indoor))) #%>%
  #select(-contains(to_remove_vec))

comb_z_df_outdoor <- comb_z_df %>%
  filter(high_speed_distance != 0)


include_mod_cols <- c("high_speed_distance", names(test_df), names(z_score_df))

comb_z_df_mod <- comb_z_df_outdoor[, names(comb_z_df_outdoor) %in% include_mod_cols]
comb_z_df_full <- comb_z_df[, names(comb_z_df) %in% include_mod_cols]

#names(comb_z_df_full)

chr_z_remove <- comb_z_df_full %>% select_if(is.character) %>%
  select(contains("_z")) %>% names()
num_z_remove <- comb_z_df_full %>% select_if(is.numeric) %>% 
  select(contains(c("binary", "position", "gender", "start_time", "end_time"))) %>%
  select(contains("z")) %>% names()

remove_vec <- c(chr_z_remove, num_z_remove)
comb_z_df_full <- comb_z_df_full %>% select(-remove_vec)
```


```{r}
# write.csv(comb_z_df_full,".\\Exported_CSVs\\comb_soc_data_cleaned_with_z_up_pos.csv",row.names=FALSE)
```




### <span style= "color:#00843D" > CLEANING LACROSSE DATA TO ENSURE `total_duration` IS INCLUDED FOR DURATION-ADJUSTED MODELING AND SAVING DATA WITH Z-SCORES </span>

##### Men's Lacrosse

Total Duration Included in CSV
```{r}
lax0_data <- read.csv(".\\Exported_CSVs\\lax0_data_cleaned.csv")
lax0_data <- lax0_data %>%
  dplyr::rename(high_speed_distance_perc=high_speed_distance, 
                high_speed_distance=high_speed_distance_covered)



not_include_indoor_vec <- c("meta", "total_effort", "heart_rate", "hr",
                            "velocity", "acceleration", "deceleration", "metre",
                            "meterage", "exertion_index","max_vel_max", 
                            "footstrikes", "running_series_count", "running_imbalance")

dur_var_names <- lax0_data %>% select(contains("duration")) %>% names()
dur_var_names <- dur_var_names <- dur_var_names[dur_var_names != "total_duration"] 
dist_var_names <- lax0_data %>% select(contains("distance")) %>% names()
dist_var_names <- dist_var_names[dist_var_names != "high_speed_distance"] 
not_include_indoor_vec <- c(not_include_indoor_vec, dur_var_names, dist_var_names)


lax0_indoor_cols_df <- lax0_data %>%
  filter(high_speed_distance > 0) %>%
  select(-contains(not_include_indoor_vec))
```

```{r}
#write.csv(lax0_indoor_cols_df,".\\Exported_CSVs\\lax0_data_cleaned_new.csv",row.names=FALSE)
```

```{r}
#lax0_data <- read.csv("LAXdata0.csv") #reading in data

lax0_data <- read.csv(".\\Exported_CSVs\\lax0_data_cleaned.csv") 

lax0_data <- lax0_data %>%
  dplyr::rename(high_speed_distance_perc = high_speed_distance, 
                high_speed_distance = high_speed_distance_covered)


not_include_indoor_vec <- c("meta", "total_effort", 
                            "heart_rate", "hr", "velocity", "acceleration",
                            "deceleration", "metre", "meterage",
                            "exertion_index","max_vel_max", 
                            "footstrikes", "running_series_count", 
                            "running_imbalance")

dur_var_names <- lax0_data %>% select(contains("duration")) %>% names()
dur_var_names <- dur_var_names <- dur_var_names[dur_var_names != "total_duration"] 
dist_var_names <- lax0_data %>% select(contains("distance")) %>% names()
dist_var_names <- dist_var_names[dist_var_names != "high_speed_distance"] 

not_include_indoor_vec <- c(not_include_indoor_vec, dur_var_names, dist_var_names)

lax0_indoor_full_df <- lax0_data %>%
  filter(high_speed_distance > 0) %>%
  select(-contains(not_include_indoor_vec))

lax0_mod_df <- lax0_indoor_full_df %>% 
  filter(total_duration > 0) 

num_col_names <- lax0_data %>% 
  select_if(is.numeric) %>% 
  names()

num_col_names_indoor <- num_col_names[num_col_names %in% names(lax0_indoor_full_df)] 

to_remove_vec <- c("unix", "position_name")

test_df <- lax0_data %>% 
  select(all_of(c("unique_session", "name_id", num_col_names_indoor))) %>%
  select(-contains(to_remove_vec))


players <- unique(lax0_data$name_id)

z_prep_mod_df <- test_df

full_df <- as.data.frame(matrix(ncol=dim(z_prep_mod_df)[2]))
names(full_df) <- names(z_prep_mod_df)

for(i in 1:length(players)){
  player_df <- z_prep_mod_df[z_prep_mod_df$name_id == players[i],]
  
  df_non_num <- player_df %>% select_if(is.character)
  df_num <- player_df %>% select_if(is.numeric)
  
  df_num <- scale(df_num)
  combined <- cbind(df_non_num, df_num)
  full_df <- rbind(full_df, combined)
}

full_df <- full_df[-1,]
names(full_df) <- paste0(names(full_df), "_z") 

z_score_df <- full_df %>% select(-contains(c("name_id")))

comb_z_df <- merge(lax0_data, z_score_df, 
                   by.x = "unique_session", by.y = "unique_session_z",
                   all.x=TRUE)

#head(comb_z_df)
```

##### PLAYER Z-SCORE CLEANING
```{r}
comb_z_df_outdoor <- comb_z_df %>%
  filter(high_speed_distance != 0)

include_mod_cols <- c("high_speed_distance", "position_name",
                      names(test_df), names(z_score_df))

comb_z_df_mod <- comb_z_df_outdoor[, names(comb_z_df_outdoor) %in% include_mod_cols]
```

```{r}
model_df <- comb_z_df_mod %>% select(!contains(c("unique_session", "name_id"))) %>%
  select("high_speed_distance", everything())

num_col_names_indoor <- num_col_names[num_col_names %in% names(lax0_indoor_full_df)]

to_remove_vec <- c("unix", "position_name_z")

test_df <- lax0_data %>%
  select(all_of(c("unique_session","name_id", "high_speed_distance", num_col_names_indoor))) %>%
  select(-contains(to_remove_vec))

comb_z_df_outdoor <- comb_z_df %>%
  filter(high_speed_distance != 0)


include_mod_cols <- c("high_speed_distance", names(test_df), names(z_score_df))

comb_z_df_mod <- comb_z_df_outdoor[, names(comb_z_df_outdoor) %in% include_mod_cols]

comb_z_df_full <- comb_z_df[, names(comb_z_df) %in% include_mod_cols]
# 
comb_z_df_full <- comb_z_df_full %>% filter(high_speed_distance >= 0)
```


```{r}
# write.csv(comb_z_df_full,".\\Exported_CSVs\\lax0_data_cleaned_with_z.csv",row.names=FALSE)
```



##### Women's Lacrosse
```{r}
lax1_data <- read.csv(".\\Exported_CSVs\\lax1_data_cleaned.csv")
lax1_data <- lax1_data %>%
  dplyr::rename(high_speed_distance=high_speed_running_distance_session)


not_include_indoor_vec <- c("meta", "total_effort", "heart_rate", "hr",
                            "velocity", "acceleration", "deceleration", "metre",
                            "meterage", "exertion_index","max_vel_max", 
                            "footstrikes", "running_series_count", "running_imbalance")

dur_var_names <- lax1_data %>% select(contains("duration")) %>% names()
dur_var_names <- dur_var_names <- dur_var_names[dur_var_names != "total_duration"] 
dist_var_names <- lax1_data %>% select(contains("distance")) %>% names()
dist_var_names <- dist_var_names[dist_var_names != "high_speed_distance"] 
not_include_indoor_vec <- c(not_include_indoor_vec, dur_var_names, dist_var_names)

lax1_indoor_cols_df <- lax1_data %>%
  #filter(high_speed_distance > 0) %>%
  select(-contains(not_include_indoor_vec))
```

```{r}
#write.csv(lax1_indoor_cols_df,".\\Exported_CSVs\\lax1_data_cleaned_new.csv",row.names=FALSE)
```

```{r}
#lax1_data <- read.csv("LAXdata1.csv") #reading in data

high_speed_dist_var_name <- "high_speed_distance"

lax1_data <- read.csv(".\\Exported_CSVs\\lax1_data_cleaned.csv") 

lax1_data <- lax1_data %>%
  dplyr::rename(high_speed_distance=high_speed_running_distance_session)


not_include_indoor_vec <- c("meta", "total_effort", 
                            "heart_rate", "hr", "velocity", "acceleration",
                            "deceleration", "metre", "meterage",
                            "exertion_index","max_vel_max", 
                            "footstrikes", "running_series_count", 
                            "running_imbalance")

dur_var_names <- lax1_data %>% select(contains("duration")) %>% names()
dur_var_names <- dur_var_names <- dur_var_names[dur_var_names != "total_duration"] 
dist_var_names <- lax1_data %>% select(contains("distance")) %>% names()
dist_var_names <- dist_var_names[dist_var_names != "high_speed_distance"] 

not_include_indoor_vec <- c(not_include_indoor_vec, dur_var_names, dist_var_names)

lax1_indoor_full_df <- lax1_data %>%
  filter(high_speed_distance > 0) %>%
  select(-contains(not_include_indoor_vec))

lax1_mod_df <- lax1_indoor_full_df %>% 
  filter(total_duration > 0) 

num_col_names <- lax1_data %>% 
  select_if(is.numeric) %>% 
  names()

num_col_names_indoor <- num_col_names[num_col_names %in% names(lax1_indoor_full_df)] 

to_remove_vec <- c("unix", "position_name")

test_df <- lax1_data %>% 
  select(all_of(c("unique_session", "name_id", num_col_names_indoor))) %>%
  select(-contains(to_remove_vec))

players <- unique(lax1_data$name_id)

z_prep_mod_df <- test_df

full_df <- as.data.frame(matrix(ncol=dim(z_prep_mod_df)[2]))
names(full_df) <- names(z_prep_mod_df)

for(i in 1:length(players)){
  player_df <- z_prep_mod_df[z_prep_mod_df$name_id == players[i],]
  
  df_non_num <- player_df %>% select_if(is.character)
  df_num <- player_df %>% select_if(is.numeric)
  
  df_num <- scale(df_num)
  combined <- cbind(df_non_num, df_num)
  full_df <- rbind(full_df, combined)
}

full_df <- full_df[-1,]
names(full_df) <- paste0(names(full_df), "_z") 

z_score_df <- full_df %>% select(-contains(c("name_id")))

comb_z_df <- merge(lax1_data, z_score_df, 
                   by.x = "unique_session", by.y = "unique_session_z",
                   all.x=TRUE)


comb_z_df_outdoor <- comb_z_df %>%
  filter(high_speed_distance != 0)

include_mod_cols <- c("high_speed_distance", "position_name",
                      names(test_df), names(z_score_df))

comb_z_df_mod <- comb_z_df_outdoor[, names(comb_z_df_outdoor) %in% include_mod_cols]
```


```{r modeling split - reg and z-score}
model_df <- comb_z_df_mod %>% select(!contains(c("unique_session", "name_id"))) %>%
  select("high_speed_distance", everything())

num_col_names_indoor <- num_col_names[num_col_names %in% names(lax1_indoor_full_df)]

to_remove_vec <- c("unix", "position_name_z", "total_duration_z")

test_df <- lax1_data %>%
  select(all_of(c("unique_session","name_id", "high_speed_distance", num_col_names_indoor))) %>%
  select(-contains(to_remove_vec))

comb_z_df_outdoor <- comb_z_df %>%
  filter(high_speed_distance != 0)

include_mod_cols <- c("high_speed_distance", names(test_df), names(z_score_df))

comb_z_df_mod <- comb_z_df_outdoor[, names(comb_z_df_outdoor) %in% include_mod_cols]

comb_z_df_full <- comb_z_df[, names(comb_z_df) %in% include_mod_cols]
# 
comb_z_df_full <- comb_z_df_full %>% filter(high_speed_distance >= 0)
```

```{r}
# write.csv(comb_z_df_full,".\\Exported_CSVs\\lax1_data_cleaned_with_z.csv",row.names=FALSE) 
```
