---
title: "PostTune_Code_032124"
author: "Kaylin Slattery"
date: "2024-03-21"
output: html_document
---
# set up 
```{r packages warnings=FALSE, error=FALSE, message=FALSE}
library(tidyverse)
library(DescTools)
library(janitor)
library(ggplot2)
library(jpeg)
# library(chron)
# library(lubridate)
# library(anytime)
library(ggpubr)
# library(devtools) 
#install_github("AppliedDataSciencePartners/xgboostExplainer")
library(xgboost) # Load XGBoost
library(caret) # Load Caret
# library(xgboostExplainer) # Load XGboost Explainer
# library(pROC) # Load proc
library(SHAPforxgboost) # Load shap for XGBoost
library(plyr)
library(Metrics)
library(kableExtra)
```

```{r nd colors}
# setting ND colors 
nd_navy <- "#0C2340"
nd_green <- "#00843D"
nd_gold <- "#C99700"
```



#  ---------------------------- DURATION WITH Z-SCORES ----------------------------
```{r}
base_path <- ".\\CRC_RDAs\\Duration"
team_dir <- "\\cs_z"
file_abr <- "\\cs_dur_"
file_names <- c("alpha_lambda.rda", "bst_final_mod.rda", 
                "eta.rda", "gamma.rda", "min_child_max_depth.rda", 
                "pre_tune_files.rda", "sample_colsample.rda")

file_names_vec <- c()

for (f in seq_along(file_names)){
  new_name <- paste0(base_path, team_dir, file_abr, file_names[f])
  file_names_vec <- c(file_names_vec, new_name)
  
  load(new_name)
}


# load(".\\CRC_RDAs\\lax1_r2\\lax1_alpha_lambda.rda")
# #res_db, bst_lambda, bst_alpha, g_5,

# load(".\\CRC_RDAs\\lax1_r2\\lax1_bst_final_mod.rda")
# # bst_final_mod, bst_pred_data, tuned_importance, imp_plot, xgb_imp_df,
# # lax1_full_preds_df, post_tune_mod_pred_viz,
# # post_tune_mod_pred_viz_2, post_tune_diff_hist

# load(".\\CRC_RDAs\\lax1_r2\\lax1_eta.rda")
# #bst_mod_1, bst_mod_2, bst_mod_3, bst_mod_4, bst_mod_5,g_6, g_7, plot_data,

# load(".\\CRC_RDAs\\lax1_r2\\lax1_gamma.rda")
# # gam_df, bst_gamma,

# load(".\\CRC_RDAs\\lax1_r2\\lax1_min_child_max_depth.rda")
# # res_db, bst_max_depth, bst_min_child, g_2, 

# load(".\\CRC_RDAs\\lax1_r2\\lax1_pre_tune_files.rda")
# # prelim_importance, bst_split_pred_data_pre_tune, pred_viz_pre_tune, diff_hist_pre_tune

# load(".\\CRC_RDAs\\lax1_r2\\lax1_sample_colsample.rda")
# # res_db, bst_col_samp, bst_sub_samp, g_4, 

tot_dur_df <- comb_dur_df %>% 
  filter(high_speed_distance > 0) %>%
  select(total_duration, high_speed_distance_per_min, high_speed_distance)
  
model_df

options(scipen = 999) # turn off scientific notation
#test_comb_tuned_df$pred_diff_meters <- test_comb_tuned_df$high_speed_distance - test_comb_tuned_df$hsd_pred

# test_comb_tuned_df %>% 
#   select(all_of(c("pred_diff_meters", "high_speed_distance", "hsd_pred")))
# summary(test_comb_tuned_df$pred_diff_meters)



options(scipen = 999) # turn off scientific notation
set.seed(33) # Set Seed
split_ratio <- 0.7
train_row_ind <- sample(x=nrow(model_df), 
                        size=floor(nrow(model_df) * split_ratio))

# Split the data into training and testing sets
train_data <- model_df[train_row_ind, ]
test_data <- model_df[-train_row_ind, ]

train_response = train_data[,1]
train_x = train_data[,-1]
test_response = test_data[,1]
test_x = test_data[,-1]

response <- model_df$high_speed_distance_per_min

# Create training data XGBOOST
dtrain <- xgb.DMatrix(data = as.matrix(train_data[ ,-1]), label = train_response)
# Create test data XGBOOST
dtest <- xgb.DMatrix(data = as.matrix(test_data[ ,-1]),label = test_response)

tot_dur_df <- comb_dur_df %>% 
  select(total_duration, high_speed_distance_per_min, high_speed_distance) %>%
  filter(high_speed_distance > 0)
#model_df[-train_row_ind, ]
tot_dur_df_test <- tot_dur_df[-train_row_ind, ]
dim(tot_dur_df_test)[1] == dim(test_data)[1]
names(tot_dur_df_test) <- c("total_duration", "actual_hsd_per_min", "actual_hsd_meters")
identical(tot_dur_df_test$actual_hsd_per_min, bst_pred_data$actual)

test_comb_tuned_df <- cbind.data.frame(bst_pred_data, tot_dur_df_test)
test_comb_tuned_df$pred_diff_meters <- test_comb_tuned_df$difference*test_comb_tuned_df$total_duration
```


```{r}
dim(test_comb_tuned_df[abs(test_comb_tuned_df$pred_diff_meters) <= 150, ])[1]/(dim(test_comb_tuned_df)[1])
xgb_imp_df[xgb_imp_df$Feature == "position_name", ] 

#writeClipboard(xgb_imp_df[1:10, ]$Feature)
```


## Post-Tune Evaluation  
### Important Variables
```{r}
# bst_final_mod, bst_pred_data, tuned_importance, imp_plot, xgb_imp_df,
# lax1_full_preds_df, post_tune_mod_pred_viz,
# post_tune_mod_pred_viz_2,post_tune_diff_hist

xgb_imp_df
ggplot(data=xgb_imp_df[1:10,], aes(x=Importance, 
                                   y=reorder(Feature, -Importance, decreasing = TRUE))) + 
  theme_minimal() + geom_col(fill=nd_green) +
  labs(title = "XG Boost Model Top 5 Most Important Variables", 
       x = "Importance", y = "Variable") 

xgb_imp_df[1:10, ]$Feature
```

### Train RMSE line
```{r}
bst_final_mod$evaluation_log %>% 
  pivot_longer(cols = c(train_rmse), names_to = "phase") %>% 
  ggplot(aes(x = iter, y = value, color = phase)) + 
  geom_line(color=nd_navy, size=2) + theme_minimal() +
  labs(title = "Train RMSE - XGBoost")
```

### Histogram Pre & Post-Tune
```{r}
# ggplot(lax1_full_preds_df, aes(x = pred_diff_meters, y=..density.., fill = time)) +
#   geom_histogram(data = subset(lax1_full_preds_df, time == "pre-tune"),
#                  #alpha = 0.5, 
#                  position = "identity", binwidth =50) +
#   # geom_histogram(data = subset(lax1_full_preds_df, time == "post-tune"),
#   #                alpha = 0.5, position = "identity", binwidth = 50) +
#   scale_fill_manual(values = c("pre-tune" = nd_navy, "post-tune" = nd_green)) +
#   labs(title = "Overlaying Histograms - lax1",
#        x = "Residual (Actual - Predicted)", y = "Frequency") +
#   theme_minimal() + xlim(-1025, 1025)

ggplot(test_comb_tuned_df,
       aes(x = pred_diff_meters, y=..density.., fill = time)) +
  geom_histogram(alpha = 0.5, position = "identity", binwidth =50) +
  # geom_histogram(data = subset(lax1_full_preds_df, time == "post-tune"),
  #                alpha = 0.5, position = "identity", binwidth = 50) +
  scale_fill_manual(values = c("pre-tune" = nd_navy, "post-tune" = nd_green)) +
  labs(title = "Histograms",
       x = "Residual (Actual - Predicted)", y = "Frequency") +
  theme_minimal() + xlim(-1500, 1500)
```

### percent of preds within distance
```{r}
# lax1_post <- lax1_full_preds_df %>% filter(time == "post-tune")
# dim(lax1_post[abs(lax1_post$pred_diff_meters) <= 150, ])[1]/(dim(lax1_post)[1]) #75.9%
# 
# lax1_pre <- lax1_full_preds_df %>% filter(time == "pre-tune")
# dim(lax1_pre[abs(lax1_pre$pred_diff_meters) <= 150, ])[1]/(dim(lax1_pre)[1])
# 
# test_comb_df
# 
# summary(lax1_pre$pred_diff_meters)

#test_comb_tuned_df$pred_diff_meters
dim(test_comb_tuned_df[abs(test_comb_tuned_df$pred_diff_meters) <= 150, ])[1]/(dim(test_comb_tuned_df)[1])

over500_diff <- dim(test_comb_tuned_df[abs(test_comb_tuned_df$pred_diff_meters) >= 500, ])[1]
over250_diff <- dim(test_comb_tuned_df[abs(test_comb_tuned_df$pred_diff_meters) >= 250, ])[1]

(over250_diff-over500_diff)/(dim(test_comb_tuned_df)[1])

# 9 % between +-250
```











# Soccer
## Loading from RDAs
```{r}
base_path <- ".\\CRC_RDAs\\Old Folders"
team_dir <- "\\soc_dur"
file_abr <- "\\cs_dur_"
file_names <- c("alpha_lambda.rda", "bst_final_mod.rda", 
                "eta.rda", "gamma.rda", "min_child_max_depth.rda", 
                "pre_tune_files.rda", "sample_colsample.rda")

file_names_vec <- c()

for (f in seq_along(file_names)){
  new_name <- paste0(base_path, team_dir, file_abr, file_names[f])
  file_names_vec <- c(file_names_vec, new_name)
  
  load(new_name)
}


# load(".\\CRC_RDAs\\lax1_r2\\lax1_alpha_lambda.rda")
# #res_db, bst_lambda, bst_alpha, g_5,

# load(".\\CRC_RDAs\\lax1_r2\\lax1_bst_final_mod.rda")
# # bst_final_mod, bst_pred_data, tuned_importance, imp_plot, xgb_imp_df,
# # lax1_full_preds_df, post_tune_mod_pred_viz,
# # post_tune_mod_pred_viz_2, post_tune_diff_hist

# load(".\\CRC_RDAs\\lax1_r2\\lax1_eta.rda")
# #bst_mod_1, bst_mod_2, bst_mod_3, bst_mod_4, bst_mod_5,g_6, g_7, plot_data,

# load(".\\CRC_RDAs\\lax1_r2\\lax1_gamma.rda")
# # gam_df, bst_gamma,

# load(".\\CRC_RDAs\\lax1_r2\\lax1_min_child_max_depth.rda")
# # res_db, bst_max_depth, bst_min_child, g_2, 

# load(".\\CRC_RDAs\\lax1_r2\\lax1_pre_tune_files.rda")
# # prelim_importance, bst_split_pred_data_pre_tune, pred_viz_pre_tune, diff_hist_pre_tune

# load(".\\CRC_RDAs\\lax1_r2\\lax1_sample_colsample.rda")
# # res_db, bst_col_samp, bst_sub_samp, g_4, 

options(scipen = 999) # turn off scientific notation
test_comb_tuned_df$pred_diff_meters <- test_comb_tuned_df$high_speed_distance - test_comb_tuned_df$hsd_pred

test_comb_tuned_df %>% 
  select(all_of(c("pred_diff_meters", "high_speed_distance", "hsd_pred")))
summary(test_comb_tuned_df$pred_diff_meters)



options(scipen = 999) # turn off scientific notation
test_comb_df$pred_diff_meters <- test_comb_df$high_speed_distance - test_comb_df$hsd_pred

test_comb_df %>% 
  select(all_of(c("pred_diff_meters", "high_speed_distance", "hsd_pred")))
summary(test_comb_df$pred_diff_meters)

#lax1_full_preds_df <- rbind.data.frame(test_comb_df, test_comb_tuned_df)

test_comb_tuned_df

```

## Post-Tune Evaluation  
### Important Variables
```{r}
# bst_final_mod, bst_pred_data, tuned_importance, imp_plot, xgb_imp_df,
# lax1_full_preds_df, post_tune_mod_pred_viz,
# post_tune_mod_pred_viz_2,post_tune_diff_hist

xgb_imp_df
ggplot(data=xgb_imp_df[1:10,], aes(x=Importance, 
                                   y=reorder(Feature, -Importance, decreasing = TRUE))) + 
  theme_minimal() + geom_col(fill=nd_green) +
  labs(title = "XG Boost Model Top 5 Most Important Variables", 
       x = "Importance", y = "Variable") 

xgb_imp_df[1:10, ]$Feature
```

### Train RMSE line
```{r}
bst_final_mod$evaluation_log %>% 
  pivot_longer(cols = c(train_rmse), names_to = "phase") %>% 
  ggplot(aes(x = iter, y = value, color = phase)) + geom_line(color=nd_navy, size=2) + theme_minimal() +
  labs(title = "Train RMSE - lax1 XGBoost")
```

### Histogram Pre & Post-Tune
```{r}
# ggplot(lax1_full_preds_df, aes(x = pred_diff_meters, y=..density.., fill = time)) +
#   geom_histogram(data = subset(lax1_full_preds_df, time == "pre-tune"),
#                  #alpha = 0.5, 
#                  position = "identity", binwidth =50) +
#   # geom_histogram(data = subset(lax1_full_preds_df, time == "post-tune"),
#   #                alpha = 0.5, position = "identity", binwidth = 50) +
#   scale_fill_manual(values = c("pre-tune" = nd_navy, "post-tune" = nd_green)) +
#   labs(title = "Overlaying Histograms - lax1",
#        x = "Residual (Actual - Predicted)", y = "Frequency") +
#   theme_minimal() + xlim(-1025, 1025)

ggplot(test_comb_tuned_df,
       aes(x = pred_diff_meters, y=..density.., fill = time)) +
  geom_histogram(alpha = 0.5, position = "identity", binwidth =50) +
  # geom_histogram(data = subset(lax1_full_preds_df, time == "post-tune"),
  #                alpha = 0.5, position = "identity", binwidth = 50) +
  scale_fill_manual(values = c("pre-tune" = nd_navy, "post-tune" = nd_green)) +
  labs(title = "Histograms - soc",
       x = "Residual (Actual - Predicted)", y = "Frequency") +
  theme_minimal() + xlim(-1025, 1025)
```

### percent of preds within distance
```{r}
# lax1_post <- lax1_full_preds_df %>% filter(time == "post-tune")
# dim(lax1_post[abs(lax1_post$pred_diff_meters) <= 150, ])[1]/(dim(lax1_post)[1]) #75.9%
# 
# lax1_pre <- lax1_full_preds_df %>% filter(time == "pre-tune")
# dim(lax1_pre[abs(lax1_pre$pred_diff_meters) <= 150, ])[1]/(dim(lax1_pre)[1])
# 
# test_comb_df
# 
# summary(lax1_pre$pred_diff_meters)

#test_comb_tuned_df$pred_diff_meters
dim(test_comb_tuned_df[abs(test_comb_tuned_df$pred_diff_meters) <= 150, ])[1]/(dim(test_comb_tuned_df)[1])

over500_diff <- dim(test_comb_tuned_df[abs(test_comb_tuned_df$pred_diff_meters) >= 500, ])[1]
over250_diff <- dim(test_comb_tuned_df[abs(test_comb_tuned_df$pred_diff_meters) >= 250, ])[1]

(over250_diff-over500_diff)/(dim(test_comb_tuned_df)[1])

# 9 % between +-250
```

### OLD Loading from RDAs
```{r}
base_path <- ".\\CRC_RDAs\\duration"
team_dir <- "\\soc"
file_abr <- "\\cs_dur_"
file_names <- c("alpha_lambda.rda", "bst_final_mod.rda", 
                "eta.rda", "gamma.rda", "min_child_max_depth.rda", 
                "pre_tune_files.rda", "sample_colsample.rda")

file_names_vec <- c()

for (f in seq_along(file_names)){
  new_name <- paste0(base_path, team_dir, file_abr, file_names[f])
  file_names_vec <- c(file_names_vec, new_name)
  
  load(new_name)
}



# load(".\\CRC_RDAs\\duration\\cs_dur_alpha_lambda.rda")
# #res_db, bst_lambda, bst_alpha, g_5,
# load(".\\CRC_RDAs\\soc_dur\\cs_dur_bst_final_mod.rda")
# # bst_final_mod, bst_pred_data, tuned_importance, imp_plot, xgb_imp_df,
# # full_preds_df, post_tune_mod_pred_viz,
# # post_tune_mod_pred_viz_2,post_tune_diff_hist
# load(".\\CRC_RDAs\\soc_dur\\cs_dur_eta.rda")
# #bst_mod_1, bst_mod_2, bst_mod_3, bst_mod_4, bst_mod_5,g_6, g_7, plot_data,
# load(".\\CRC_RDAs\\soc_dur\\cs_dur_gamma.rda")
# # gam_df, bst_gamma,
# load(".\\CRC_RDAs\\soc_dur\\cs_dur_min_child_max_depth.rda")
# # res_db, bst_max_depth, bst_min_child, g_2, 
# load(".\\CRC_RDAs\\soc_dur\\cs_dur_pre_tune_files.rda")
# # prelim_importance, bst_split_pred_data_pre_tune, pred_viz_pre_tune, diff_hist_pre_tune
# load(".\\CRC_RDAs\\soc_dur\\cs_dur_sample_colsample.rda")
# # res_db, bst_col_samp, bst_sub_samp, g_4, 


bst_split_pred_data_pre_tune
bst_pred_data
#full_pred_soc_df <- rbind.data.frame(bst_split_pred_data_pre_tune, bst_pred_data)
#test_comb_df

options(scipen = 999) # turn off scientific notation
test_comb_tuned_df$pred_diff_meters <- test_comb_tuned_df$high_speed_distance - test_comb_tuned_df$hsd_pred

test_comb_tuned_df %>% 
  select(all_of(c("pred_diff_meters", "high_speed_distance", "hsd_pred")))
summary(test_comb_tuned_df$pred_diff_meters)
# predicting for duration
```

## Post-Tune Evaluation  
### Important Variables
```{r}
# bst_final_mod, bst_pred_data, tuned_importance, imp_plot, xgb_imp_df,
# full_preds_df, post_tune_mod_pred_viz,
# post_tune_mod_pred_viz_2,post_tune_diff_hist

xgb_imp_df
ggplot(data=xgb_imp_df[1:10,], aes(x=Importance, 
y=reorder(Feature, -Importance, decreasing = TRUE))) + 
  theme_minimal() + geom_col(fill=nd_navy) +
  labs(title = "Soccer XGBoost Model Top 10 Most Important Variables", 
       x = "Importance", y = "Variable") 

#xgb_imp_df[1:10, ]$Feature

xgb_imp_df$z_var_bin <- ifelse(grepl("z", xgb_imp_df$Feature), "z-score", "non-z")

# Now plot using this new column for fill
ggplot(data=xgb_imp_df[1:10, ], aes(x=Importance,
                                    y=reorder(Feature, -Importance, decreasing = TRUE))) +
  theme_minimal() +
  geom_col(aes(fill=z_var_bin)) +  
  scale_fill_manual(values=c("z-score"=nd_green, "non-z"=nd_navy)) +
  labs(title="Top Predictors - Soccer XGBoost", x="Importance", y="Variable")
```

### Train RMSE line
```{r}
bst_final_mod$evaluation_log %>% 
  pivot_longer(cols = c(train_rmse), names_to = "phase") %>% 
  ggplot(aes(x = iter, y = value, color = phase)) + geom_line(color=nd_navy, size=2) + theme_minimal() +
  labs(title = "Train RMSE - Soccer XGBoost")
```

### Histogram Pre & Post Tune 
```{r}
ggplot(full_pred_soc_df, aes(x = difference, y=..density.., fill = time)) + 
  geom_histogram(data = subset(full_pred_soc_df, time == "pre-tune"),
                 alpha = 0.5, position = "identity", binwidth =50) + 
  geom_histogram(data = subset(full_pred_soc_df, time == "post-tune"),
                 alpha = 0.5, position = "identity", binwidth = 50) + 
  scale_fill_manual(values = c("pre-tune" = nd_navy, "post-tune" = nd_green)) + 
  labs(title = "Overlaying Histograms - Soccer",
       x = "Residual (Actual - Predicted)", y = "Frequency") + 
  theme_minimal() + xlim(-1500, 1500)
```

```{r}
post <- full_pred_soc_df %>% filter(time == "post-tune")
dim(post[abs(post$) <= 150, ])[1]/(dim(post)[1]) #75.9%
```

# Women's Lacrosse
## Loading from RDAs
```{r}
base_path <- ".\\CRC_RDAs\\Old Folders"
team_dir <- "\\lax1_dur_r3"
file_abr <- "\\l1_dur_"
file_names <- c("alpha_lambda.rda", "bst_final_mod.rda", 
                "eta.rda", "gamma.rda", "min_child_max_depth.rda", 
                "pre_tune_files.rda", "sample_colsample.rda")

file_names_vec <- c()

for (f in seq_along(file_names)){
  new_name <- paste0(base_path, team_dir, file_abr, file_names[f])
  file_names_vec <- c(file_names_vec, new_name)
  
  load(new_name)
}


# load(".\\CRC_RDAs\\lax1_r2\\lax1_alpha_lambda.rda")
# #res_db, bst_lambda, bst_alpha, g_5,

# load(".\\CRC_RDAs\\lax1_r2\\lax1_bst_final_mod.rda")
# # bst_final_mod, bst_pred_data, tuned_importance, imp_plot, xgb_imp_df,
# # lax1_full_preds_df, post_tune_mod_pred_viz,
# # post_tune_mod_pred_viz_2, post_tune_diff_hist

# load(".\\CRC_RDAs\\lax1_r2\\lax1_eta.rda")
# #bst_mod_1, bst_mod_2, bst_mod_3, bst_mod_4, bst_mod_5,g_6, g_7, plot_data,

# load(".\\CRC_RDAs\\lax1_r2\\lax1_gamma.rda")
# # gam_df, bst_gamma,

# load(".\\CRC_RDAs\\lax1_r2\\lax1_min_child_max_depth.rda")
# # res_db, bst_max_depth, bst_min_child, g_2, 

# load(".\\CRC_RDAs\\lax1_r2\\lax1_pre_tune_files.rda")
# # prelim_importance, bst_split_pred_data_pre_tune, pred_viz_pre_tune, diff_hist_pre_tune

# load(".\\CRC_RDAs\\lax1_r2\\lax1_sample_colsample.rda")
# # res_db, bst_col_samp, bst_sub_samp, g_4, 

options(scipen = 999) # turn off scientific notation
test_comb_tuned_df$pred_diff_meters <- test_comb_tuned_df$high_speed_distance - test_comb_tuned_df$hsd_pred

test_comb_tuned_df %>% 
  select(all_of(c("pred_diff_meters", "high_speed_distance", "hsd_pred")))
summary(test_comb_tuned_df$pred_diff_meters)



options(scipen = 999) # turn off scientific notation
test_comb_df$pred_diff_meters <- test_comb_df$high_speed_distance - test_comb_df$hsd_pred

test_comb_df %>% 
  select(all_of(c("pred_diff_meters", "high_speed_distance", "hsd_pred")))
summary(test_comb_df$pred_diff_meters)

lax1_full_preds_df <- rbind.data.frame(test_comb_df, test_comb_tuned_df)

test_comb_tuned_df

```

## Post-Tune Evaluation  
### Important Variables
```{r}
# bst_final_mod, bst_pred_data, tuned_importance, imp_plot, xgb_imp_df,
# lax1_full_preds_df, post_tune_mod_pred_viz,
# post_tune_mod_pred_viz_2,post_tune_diff_hist

xgb_imp_df
ggplot(data=xgb_imp_df[1:10,], aes(x=Importance, 
                                   y=reorder(Feature, -Importance, decreasing = TRUE))) + 
  theme_minimal() + geom_col(fill=nd_green) +
  labs(title = "XG Boost Model Top 5 Most Important Variables", 
       x = "Importance", y = "Variable") 

xgb_imp_df[1:10, ]$Feature
```

### Train RMSE line
```{r}
bst_final_mod$evaluation_log %>% 
  pivot_longer(cols = c(train_rmse), names_to = "phase") %>% 
  ggplot(aes(x = iter, y = value, color = phase)) + geom_line(color=nd_navy, size=2) + theme_minimal() +
  labs(title = "Train RMSE - lax1 XGBoost")
```

### Histogram Pre & Post-Tune
```{r}
# ggplot(lax1_full_preds_df, aes(x = pred_diff_meters, y=..density.., fill = time)) +
#   geom_histogram(data = subset(lax1_full_preds_df, time == "pre-tune"),
#                  #alpha = 0.5, 
#                  position = "identity", binwidth =50) +
#   # geom_histogram(data = subset(lax1_full_preds_df, time == "post-tune"),
#   #                alpha = 0.5, position = "identity", binwidth = 50) +
#   scale_fill_manual(values = c("pre-tune" = nd_navy, "post-tune" = nd_green)) +
#   labs(title = "Overlaying Histograms - lax1",
#        x = "Residual (Actual - Predicted)", y = "Frequency") +
#   theme_minimal() + xlim(-1025, 1025)

ggplot(test_comb_tuned_df,
       aes(x = pred_diff_meters, y=..density.., fill = time)) +
  geom_histogram(alpha = 0.5, position = "identity", binwidth =50) +
  # geom_histogram(data = subset(lax1_full_preds_df, time == "post-tune"),
  #                alpha = 0.5, position = "identity", binwidth = 50) +
  scale_fill_manual(values = c("pre-tune" = nd_navy, "post-tune" = nd_green)) +
  labs(title = "Histograms - lax1",
       x = "Residual (Actual - Predicted)", y = "Frequency") +
  theme_minimal() + xlim(-1025, 1025)
```

### percent of preds within distance
```{r}
# lax1_post <- lax1_full_preds_df %>% filter(time == "post-tune")
# dim(lax1_post[abs(lax1_post$pred_diff_meters) <= 150, ])[1]/(dim(lax1_post)[1]) #75.9%
# 
# lax1_pre <- lax1_full_preds_df %>% filter(time == "pre-tune")
# dim(lax1_pre[abs(lax1_pre$pred_diff_meters) <= 150, ])[1]/(dim(lax1_pre)[1])
# 
# test_comb_df
# 
# summary(lax1_pre$pred_diff_meters)

test_comb_tuned_df$pred_diff_meters
dim(test_comb_tuned_df[abs(test_comb_tuned_df$pred_diff_meters) <= 150, ])[1]/(dim(test_comb_tuned_df)[1])
```



# Men's Lacrosse
## Loading from RDAs
```{r}
base_path <- ".\\CRC_RDAs\\Old Folders"
team_dir <- "\\lax0_dur_r1"
file_abr <- "\\l0_dur_"
file_names <- c("alpha_lambda.rda", "bst_final_mod.rda", 
                "eta.rda", "gamma.rda", "min_child_max_depth.rda", 
                "pre_tune_files.rda", "sample_colsample.rda")

file_names_vec <- c()

for (f in seq_along(file_names)){
  new_name <- paste0(base_path, team_dir, file_abr, file_names[f])
  file_names_vec <- c(file_names_vec, new_name)
  
  load(new_name)
}


# load(".\\CRC_RDAs\\lax1_r2\\lax1_alpha_lambda.rda")
# #res_db, bst_lambda, bst_alpha, g_5,

# load(".\\CRC_RDAs\\lax1_r2\\lax1_bst_final_mod.rda")
# # bst_final_mod, bst_pred_data, tuned_importance, imp_plot, xgb_imp_df,
# # lax1_full_preds_df, post_tune_mod_pred_viz,
# # post_tune_mod_pred_viz_2, post_tune_diff_hist

# load(".\\CRC_RDAs\\lax1_r2\\lax1_eta.rda")
# #bst_mod_1, bst_mod_2, bst_mod_3, bst_mod_4, bst_mod_5,g_6, g_7, plot_data,

# load(".\\CRC_RDAs\\lax1_r2\\lax1_gamma.rda")
# # gam_df, bst_gamma,

# load(".\\CRC_RDAs\\lax1_r2\\lax1_min_child_max_depth.rda")
# # res_db, bst_max_depth, bst_min_child, g_2, 

# load(".\\CRC_RDAs\\lax1_r2\\lax1_pre_tune_files.rda")
# # prelim_importance, bst_split_pred_data_pre_tune, pred_viz_pre_tune, diff_hist_pre_tune

# load(".\\CRC_RDAs\\lax1_r2\\lax1_sample_colsample.rda")
# # res_db, bst_col_samp, bst_sub_samp, g_4, 

options(scipen = 999) # turn off scientific notation
test_comb_tuned_df$pred_diff_meters <- test_comb_tuned_df$high_speed_distance - test_comb_tuned_df$hsd_pred

test_comb_tuned_df %>% 
  select(all_of(c("pred_diff_meters", "high_speed_distance", "hsd_pred")))
summary(test_comb_tuned_df$pred_diff_meters)



options(scipen = 999) # turn off scientific notation
test_comb_df$pred_diff_meters <- test_comb_df$high_speed_distance - test_comb_df$hsd_pred

test_comb_df %>% 
  select(all_of(c("pred_diff_meters", "high_speed_distance", "hsd_pred")))
summary(test_comb_df$pred_diff_meters)

#lax1_full_preds_df <- rbind.data.frame(test_comb_df, test_comb_tuned_df)

test_comb_tuned_df

```

## Post-Tune Evaluation  
### Important Variables
```{r}
# bst_final_mod, bst_pred_data, tuned_importance, imp_plot, xgb_imp_df,
# lax1_full_preds_df, post_tune_mod_pred_viz,
# post_tune_mod_pred_viz_2,post_tune_diff_hist

xgb_imp_df
ggplot(data=xgb_imp_df[1:10,], aes(x=Importance, 
                                   y=reorder(Feature, -Importance, decreasing = TRUE))) + 
  theme_minimal() + geom_col(fill=nd_green) +
  labs(title = "XG Boost Model Top 5 Most Important Variables", 
       x = "Importance", y = "Variable") 

xgb_imp_df[1:10, ]$Feature
```

### Train RMSE line
```{r}
bst_final_mod$evaluation_log %>% 
  pivot_longer(cols = c(train_rmse), names_to = "phase") %>% 
  ggplot(aes(x = iter, y = value, color = phase)) + geom_line(color=nd_navy, size=2) + theme_minimal() +
  labs(title = "Train RMSE - lax1 XGBoost")
```

### Histogram Pre & Post-Tune
```{r}
# ggplot(lax1_full_preds_df, aes(x = pred_diff_meters, y=..density.., fill = time)) +
#   geom_histogram(data = subset(lax1_full_preds_df, time == "pre-tune"),
#                  #alpha = 0.5, 
#                  position = "identity", binwidth =50) +
#   # geom_histogram(data = subset(lax1_full_preds_df, time == "post-tune"),
#   #                alpha = 0.5, position = "identity", binwidth = 50) +
#   scale_fill_manual(values = c("pre-tune" = nd_navy, "post-tune" = nd_green)) +
#   labs(title = "Overlaying Histograms - lax1",
#        x = "Residual (Actual - Predicted)", y = "Frequency") +
#   theme_minimal() + xlim(-1025, 1025)

ggplot(test_comb_tuned_df,
       aes(x = pred_diff_meters, y=..density.., fill = time)) +
  geom_histogram(alpha = 0.5, position = "identity", binwidth =50) +
  # geom_histogram(data = subset(lax1_full_preds_df, time == "post-tune"),
  #                alpha = 0.5, position = "identity", binwidth = 50) +
  scale_fill_manual(values = c("pre-tune" = nd_navy, "post-tune" = nd_green)) +
  labs(title = "Histograms - lax0",
       x = "Residual (Actual - Predicted)", y = "Frequency") +
  theme_minimal() + xlim(-1025, 1025)
```

### percent of preds within distance
```{r}
# lax1_post <- lax1_full_preds_df %>% filter(time == "post-tune")
# dim(lax1_post[abs(lax1_post$pred_diff_meters) <= 150, ])[1]/(dim(lax1_post)[1]) #75.9%
# 
# lax1_pre <- lax1_full_preds_df %>% filter(time == "pre-tune")
# dim(lax1_pre[abs(lax1_pre$pred_diff_meters) <= 150, ])[1]/(dim(lax1_pre)[1])
# 
# test_comb_df
# 
# summary(lax1_pre$pred_diff_meters)

#test_comb_tuned_df$pred_diff_meters
dim(test_comb_tuned_df[abs(test_comb_tuned_df$pred_diff_meters) <= 150, ])[1]/(dim(test_comb_tuned_df)[1])

over500_diff <- dim(test_comb_tuned_df[abs(test_comb_tuned_df$pred_diff_meters) >= 500, ])[1]
over250_diff <- dim(test_comb_tuned_df[abs(test_comb_tuned_df$pred_diff_meters) >= 250, ])[1]

(over250_diff-over500_diff)/(dim(test_comb_tuned_df)[1])

# 9 % between +-250
```



### Loading from RDAs
```{r}
load(".\\CRC_RDAs\\lax0_r2\\lax0_alpha_lambda.rda")
#res_db, bst_lambda, bst_alpha, g_5,
load(".\\CRC_RDAs\\lax0_r2\\lax0_bst_final_mod.rda")
# bst_final_mod, bst_pred_data, tuned_importance, imp_plot, xgb_imp_df,
# lax0_full_preds_df, post_tune_mod_pred_viz,
# post_tune_mod_pred_viz_2,post_tune_diff_hist
load(".\\CRC_RDAs\\lax0_r2\\lax0_eta.rda")
#bst_mod_1, bst_mod_2, bst_mod_3, bst_mod_4, bst_mod_5,g_6, g_7, plot_data,
load(".\\CRC_RDAs\\lax0_r2\\lax0_gamma.rda")
# gam_df, bst_gamma,
load(".\\CRC_RDAs\\lax0_r2\\lax0_min_child_max_depth.rda")
# res_db, bst_max_depth, bst_min_child, g_2, 
load(".\\CRC_RDAs\\lax0_r2\\lax0_pre_tune_files.rda")
# prelim_importance, bst_split_pred_data_pre_tune, pred_viz_pre_tune, diff_hist_pre_tune
load(".\\CRC_RDAs\\lax0_r2\\lax0_sample_colsample.rda")
# res_db, bst_col_samp, bst_sub_samp, g_4, 
```

## Post-Tune Evaluation
### Important Variables
```{r}
# bst_final_mod, bst_pred_data, tuned_importance, imp_plot, xgb_imp_df,
# lax0_full_preds_df, post_tune_mod_pred_viz,
# post_tune_mod_pred_viz_2,post_tune_diff_hist

xgb_imp_df
ggplot(data=xgb_imp_df[1:10,], aes(x=Importance, 
                                                       y=reorder(Feature, -Importance, decreasing = TRUE))) + 
  theme_minimal() + geom_col(fill=nd_navy) +
  labs(title = "XG Boost Model Top 5 Most Important Variables", 
       x = "Importance", y = "Variable") 

mlax_imp_10 <- xgb_imp_df[1:10, ]$Feature

wlax_imp_10 <- c(
   "player_load_band_3_total_player_load_z"           , "player_load_band_2_total_player_load_z"           
, "edi_z"                                            
, "average_player_load_slow_session_z"               
, "player_load_slow_z"                               
, "pl_hi_sess_z"                                     
, "ima_7_o_clock_low_1_0_z"                          
,"player_load_1d_up_z"                              
, "player_load_band_2_average_effort_count_session_z"
, "average_player_load_slow_z" 
)

wlax_imp_10[wlax_imp_10 %in% mlax_imp_10]
```

### Train RMSE Line
```{r}
bst_final_mod$evaluation_log %>% 
  pivot_longer(cols = c(train_rmse), names_to = "phase") %>% 
  ggplot(aes(x = iter, y = value, color = phase)) + geom_line(color=nd_navy, size=2) + theme_minimal() +
  labs(title = "Train RMSE - Lax0 XGBoost")

post_df <- lax0_full_preds_df %>% 
  filter(time == "post-tune") 

summary(post_df$difference)
```


### Histogram Pre & Post-Tune
```{r}
ggplot(lax0_full_preds_df, aes(x = difference, y=..density.., fill = time)) + 
  geom_histogram(data = subset(lax0_full_preds_df, time == "pre-tune"),
                 alpha = 0.5, position = "identity", binwidth =50) + 
  geom_histogram(data = subset(lax0_full_preds_df, time == "post-tune"),
                 alpha = 0.5, position = "identity", binwidth = 50) + 
  scale_fill_manual(values = c("pre-tune" = nd_navy, "post-tune" = nd_green)) + 
  labs(title = "Overlaying Histograms - Lax0",
       x = "Residual (Actual - Predicted)", y = "Frequency") + 
  theme_minimal() 
```

### Percent of Preds within distnce
```{r}
lax0_post <- lax0_full_preds_df %>% filter(time == "post-tune")
dim(lax0_post[abs(lax0_post$difference) <= 150, ])[1]/(dim(lax0_post)[1]) #75.9%
```